---
title: "UN Gender Stats Explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: https://github.com/mine-cetinkaya-rundel/rladies-phl-shiny/blob/master/01-flexdash/un-women-dash.Rmd
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(DT)
```

```{r load-data, include = FALSE}
unwomen <- read_csv("data/un-women.csv") %>% 
  dplyr::filter(
    !is.na(region),
    sex == "Female"
    ) %>%
  select(-key)
```

```{r save-fields, include = FALSE}
min_year  <- min(unwomen$year)
max_year  <- max(unwomen$year)
n_tot_years <- (max_year - min_year) + 1
regions   <- unwomen %>% distinct(region) %>% arrange(region) %>% pull(region)
```

Sidebar {.sidebar}
=======================================================================

### UN Gender Stats Data

Filter the data for characteristics you're interested in exploring:

```{r}
hr()

sliderInput(inputId = "year", label = "Year",
            min = min_year, max = max_year,
            value = c(2001, max_year), step = 1, sep = "")

hr()

checkboxGroupInput(inputId = "region", label = "Region",
                   choices = regions, 
                   selected = c(
                     "Developed regions", 
                     "Latin America and the Caribbean",
                     "Sub-Saharan Africa")
                   )

hr()

x_var <- reactive({ syms(input$x) })
y_var <- reactive({ syms(input$x) })

sel_data <- reactive({
  unwomen %>%
    dplyr::filter(
      between(year, input$year[1], input$year[2]),
      region %in% input$region
      ) %>%
    dplyr::filter(!is.na( !!!x_var() ), !is.na( !!!y_var() )) %>%
    group_by(country) %>%
    arrange(desc(year)) %>%
    top_n(1)
})
```

Dashboard
=======================================================================

Row
-----------------------------------------------------------------------

### Percentage of countries with data on selected variables {.value-box}

```{r}
renderGauge({
  
  n_sel_country <- unwomen %>%
    dplyr::filter(region %in% input$region) %>%
    distinct(country) %>%
    nrow()
  
  p <- round( nrow(sel_data()) / n_sel_country * 100, 0)
  
  gauge(p, min = 0, max = 100, symbol = "%", 
        gaugeSectors(success = c(80, 100), warning = c(40, 79), 
                     danger = c(0, 39)))
})
```

### Average percentage of women judges {.value-box}

```{r}
renderValueBox({
  valueBox(value = round(mean(sel_data()$prop_female_judge, na.rm = TRUE), 2), 
           icon = "glyphicon-bullhorn")
})
```

### Average percentage of women in parliament {.value-box}

```{r}
renderValueBox({
  valueBox(value = round(mean(sel_data()$prop_nat_parl, na.rm = TRUE), 2), 
           icon = "glyphicon-thumbs-down")
})
```

Row
-----------------------------------------------------------------------

### Variable choices {data-width=340}

Select the variables you would like to display on the X and Y axes of 
the scatterplot on the right, respectively.

```{r}
selectInput(
  inputId = "x", label = "X-axis",
  choices = c(
    "Average number of hours spent on unpaid domestic and care work" 
      = "hrs_unpaid_dom_care_work",
    "Average number of hours spent on paid and unpaid domestic and care work combined"
      = "hrs_dom_care_work",
    "Labour force participation rate" = "labor_force",
    "Proportion of employed who are self-employed" = "prop_self_employed",
    "Proportion of employed who are contributing family workers" = "prop_family_worker",
    "Proportion of employed who are employers" = "prop_employers",
    "Percentage of employed in industrial sector" = "perc_industrial",
    "Percentage of employed in service sector" = "perc_service",
    "Proportion using the Internet" = "prop_internet",
    "Proportion of seats held by women in national parliament" = "prop_nat_parl",
    "Proportion of female judges" = "prop_female_judge"
    ),
  selected = "labor_force"
  )

selectInput(
  inputId = "y", label = "Y-axis",
  choices = c(
    "Average number of hours spent on unpaid domestic and care work" 
      = "hrs_unpaid_dom_care_work",
    "Average number of hours spent on paid and unpaid domestic and care work combined"
      = "hrs_dom_care_work",
    "Labour force participation rate" = "labor_force",
    "Proportion of employed who are self-employed" = "prop_self_employed",
    "Proportion of employed who are contributing family workers" = "prop_family_worker",
    "Proportion of employed who are employers" = "prop_employers",
    "Percentage of employed in industrial sector" = "perc_industrial",
    "Percentage of employed in service sector" = "perc_service",
    "Proportion using the Internet" = "prop_internet",
    "Proportion of seats held by women in national parliament" = "prop_nat_parl",
    "Proportion of female judges" = "prop_female_judge"
    ),
  selected = "prop_female_judge"
  )

xlab <- reactive({
  case_when(
    input$x == "hrs_unpaid_dom_care_work" ~  
      "Average number of hours spent on unpaid domestic and care work",
    input$x == "hrs_dom_care_work"        ~  
      "Average number of hours spent on paid and unpaid domestic and care work combined",
    input$x == "labor_force"              ~ 
      "Labour force participation rate, ages 15-24",
    input$x == "prop_self_employed"       ~ "Proportion of employed who are self-employed",
    input$x == "prop_family_worker"       ~ 
      "Proportion of employed who are contributing family workers",
    input$x == "prop_employers"           ~ "Proportion of employed who are employers",
    input$x == "perc_industrial"          ~ "Percentage of employed in industrial sector",
    input$x == "perc_service"             ~ "Percentage of employed in service sector",
    input$x == "prop_internet"            ~ "Proportion using the Internet",
    input$x == "prop_nat_parl"            ~ 
      "Proportion of seats held by women in national parliament",
    input$x == "prop_female_judge"        ~ "Proportion of female judges"
  )
})

ylab <- reactive({
  case_when(
    input$y == "hrs_unpaid_dom_care_work" ~  
      "Average number of hours spent on unpaid domestic and care work",
    input$y == "hrs_dom_care_work"        ~  
      "Average number of hours spent on paid and unpaid domestic and care work combined",
    input$y == "labor_force"              ~ 
      "Labour force participation rate, ages 15-24",
    input$y == "prop_self_employed"       ~ "Proportion of employed who are self-employed",
    input$y == "prop_family_worker"       ~ 
      "Proportion of employed who are contributing family workers",
    input$y == "prop_employers"           ~ "Proportion of employed who are employers",
    input$y == "perc_industrial"          ~ "Percentage of employed in industrial sector",
    input$y == "perc_service"             ~ "Percentage of employed in service sector",
    input$y == "prop_internet"            ~ "Proportion using the Internet",
    input$y == "prop_nat_parl"            ~ 
      "Proportion of seats held by women in national parliament",
    input$y == "prop_female_judge"        ~ "Proportion of female judges"
  )
})

br()
```

**Note:** Only data from the most recent year for your selected variables will 
be plotted. For example, you selected years 2001 through 2016 on the left, and 
want to visualize labour force vs. proportion in the national parliament, the 
app first determines the most recent year for which there is data recorded on 
a given country in your selected regions, and visualizes only those data points.

### Visualization {data-width=700}

```{r}
renderPlot({
  ggplot(data = sel_data(), 
         mapping = aes_string(x = input$x, y = input$y, color = "region")) +
    geom_point(size = 2, alpha = 0.8) +
    theme_minimal() +
    labs(x = xlab(), y = ylab(), color = "Region")
  })
```

Data
=======================================================================

### About the data

This [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) uses data 
from [UN Gender Stats](https://genderstats.un.org).

The data used in the analysis is shown below.

```{r}
DT::renderDataTable({
  datatable( sel_data() )
})
```

